---
title: "Lab Assignment DUE 9/24"
output: html_document
---

Lab Assignment #2 - you've got this!! In this assignment, you'll practice some of the skills we learned in class on 9/17, including grouping/summarizing data, filtering data, and merging data sources.

It's okay for you to Google or use Chat-GPT for help on specific questions. But remember that this is graded for **EFFORT**, not accuracy. I'd rather you attempt something on your own and it be wrong than it be perfectly correct but you have no clue what it means.

First prove to me that the number 24 is in (hint: **%in%**) the range of numbers from 20 to 30.

```{r}

```

**Load** in the package that you need to flexibly work with data. (Hint: It's the only package we've been using.)

```{r}

```

**Bring in** **and save** the csv file called `CCRB_officers.csv`.

```{r}

```

**Make a new column** that indicates the percentage of complaints that are substantiated for each officer.

```{r}

```

Excluding officers with no complaints, what is the **average** % substantiated complaints across all officers?

```{r}

```

Now, show me the **average number of complaints for each gender**. (Hint: `group_by()` and `summarize()` will come in handy here.)

```{r}

```

**Make a new variable** that binarizes the rank variable, and simply tells us whether the rank is police officer, or higher than police officer.

```{r}

```

Bring in the `CCRB_penalties.csv` dataset.

```{r}

```

**Combine** the two dataframes into one, **joining** by the Tax_ID variable.

```{r}

```

You're done!

To show me you've finished, click the Knit button above. That will create either an HTML or PDF version of this document, and will save it in your project folder.
