---
title: "Lab Assignment DUE 10/29"
output: html_document
---

Welcome to your homework. Time to do some descriptive and inferential statistics!!

For this homework assignment, we'll be sticking with the education theme and working with a dataset on Pell grants. In case you don't know, Pell grants are one of the main ways the federal government provides financial aid for higher education to low-income students. The number of Pell grants that a school receives can often be a good indication of the number of low-income students at that school.

This data only goes up to 2017 (starting in 1999), so even if you yourself have Pell grants, you aren't included in this dataset.

This data is from [here](https://github.com/rfordatascience/tidytuesday/tree/main/data/2022/2022-08-30), where you can learn more about what each column means.

### Question 1

Let's load in `tidyverse` and read in `pell.csv`.

```{r}

```

### Question 2

Create a new dataframe that tells us the average number of Pell grant recipients in each year, in each state.

```{r}

```

### Question 3

Create another summary dataframe that tells us, for each school in NY, the average amount of Pell money a school has received (Hint: This means averaging over all years in the dataset), as well as the standard deviation of award amount.

```{r}

```

### Question 4

Plot a histogram of the number of Pell recipients at Columbia. (There are many schools across the country with Columbia in the name, so be specific when filtering.)

```{r}

```

### Question 5

Let's dive into some inferential statistics.

Using formula notation, run an independent samples t-test comparing the number of Pell grant recipients in NY to the number of Pell grant recipients in another state - your choice!

```{r}

```

### Question 6

Now run another independent samples t-test, again using formula notation, comparing the amount of Pell grant dollars in a state of your choice in 2000 to the amount of Pell grant dollars in that same state in 2010.

```{r}

```

### Question 7

Run a linear regression that models how the number of Pell grant recipients changes over time at Columbia.

```{r}

```

### Question 8

In the space below, explain what your linear regression shows in plain language. Is there an effect of time? In which direction? How do you know?

### Question 9

Let's bring it all together. Run a linear regression with two predictors: year and state. Your DV is number of Pell grant recipients. For state, you're comparing between two states. You can pick which states you'd like to compare.

```{r}

```

### Question 10

Again, explain your results. Are there any main effects or interactions? How do you know?
