---
title: "Lab Assignment DUE 9/17"
output:
  html_document: default
  pdf_document: default
date: "2025-09-09"
---

Welcome to your first lab assignment!!!
Are you excited???

For this first assignment, you're going to practice some of the **coding basics** you learned in class on 9/10.

It's okay for you to Google or use Chat-GPT for help on specific questions.
But remember that this is graded for **EFFORT**, not accuracy.
I'd rather you attempt something on your own and it be wrong than it be perfectly correct but you have no clue what it means.

If you're struggling to figure out the answer to a question, **feel free to insert a comment** letting me know what's tripping you up!
If something you do gives you an error, don't be afraid to read the error and try to figure out what it's telling you.

First, let's test some knowledge of basic R functionality.
**Create a variable** with a value of 6.
Then, using that variable, **create another variable** that is equal to 24.

```{r}

```

Now **create a variable** equal to the name of your hometown.
**Display** the number of characters in the name.

```{r}

```

**Create another variable** that is equal to your own name.
**Divide** the length of your name by the length of your hometown name, *without using any numbers in your code.*

```{r}

```

Okay, enough nonsense.
Let's work with some real data!

First, **load** the `tidyverse` package.

```{r}

```

Now **load** in the data file that has incarceration trends for New York state.
Make sure to **assign** it an object that you can re-use.
Give the variable a name that describes the data.

```{r}

```

This file has historical data on incarceration in NYS in both state prisons and county jails.

Using indexing, **output** the value of the *jail* population in 2012.

```{r}

```

**Create a new column** in your dataframe that displays the rate of total *prison* incarceration per 100,000 people, for each year.
This is a common way of reporting incarceration rates.
(*Hint: In class we multiplied the rate by 100. Here, you would multiple it by 100,000.*) How does the rate change over time?

```{r}

```

Now **create a new variable** that says what percentage of the prison population is male.

```{r}

```

What is the **average** of that percentage across all years in the dataset?
(*Hint: Are there any blanks in the dataset? How do you deal with that?*)

```{r}

```

Do the same thing - calculate average % of incarcerated people who are male - for the jail population.
Then, **display** the average.
Is it higher or lower than it was for the prison population?

```{r}

```

You're done!

To show me you've finished, click the Knit button above.
That will create either an HTML or PDF version of this document, and will save it in your project folder.
